---
title: "Project"
: Haena Kim
---

Initial data cleaning
```{r}
library(dplyr)
library(randomForest)
library(corrplot)
library(MASS)
library(car)
library(caret)

setwd("C:\\Users\\haenakim\\Documents\\GIT\\INDE498_Project")

train <- read.csv("train.csv")
test <- read.csv("test.csv")

train

# finding the dimension of our datasets
dim(train) 

#Data cleaning for train
train$Alley <- as.character(train$Alley)
train$Alley[is.na(train$Alley)] <- "none"
train$Alley <- as.factor(train$Alley)

train$MasVnrType <- as.character(train$MasVnrType)
train$MasVnrType[is.na(train$MasVnrType)] <- "none"
train$MasVnrType <- as.factor(train$MasVnrType)

train$MasVnrArea[is.na(train$MasVnrArea)] <- 0

train$BsmtQual <- as.character(train$BsmtQual)
train$BsmtQual[is.na(train$BsmtQual)] <- "none"
train$BsmtQual <- as.factor(train$BsmtQual)

train$BsmtCond <- as.character(train$BsmtCond)
train$BsmtCond[is.na(train$BsmtCond)] <- "none"
train$BsmtCond <- as.factor(train$BsmtCond)

train$BsmtExposure <- as.character(train$BsmtExposure)
train$BsmtExposure[is.na(train$BsmtExposure)] <- "none"
train$BsmtExposure <- as.factor(train$BsmtExposure)

train$BsmtFinType1 <- as.character(train$BsmtFinType1)
train$BsmtFinType1[is.na(train$BsmtFinType1)] <- "none"
train$BsmtFinType1 <- as.factor(train$BsmtFinType1)

train$BsmtFinType2 <- as.character(train$BsmtFinType2)
train$BsmtFinType2[is.na(train$BsmtFinType2)] <- "none"
train$BsmtFinType2 <- as.factor(train$BsmtFinType2)

train$FireplaceQu <- as.character(train$FireplaceQu)
train$FireplaceQu[is.na(train$FireplaceQu)] <- "none"
train$FireplaceQu <- as.factor(train$FireplaceQu)

train$GarageType <- as.character(train$GarageType)
train$GarageType[is.na(train$GarageType)] <- "none"
train$GarageType <- as.factor(train$GarageType)

train$GarageFinish <- as.character(train$GarageFinish)
train$GarageFinish[is.na(train$GarageFinish)] <- "none"
train$GarageFinish <- as.factor(train$GarageFinish)

train$GarageQual <- as.character(train$GarageQual)
train$GarageQual[is.na(train$GarageQual)] <- "none"
train$GarageQual <- as.factor(train$GarageQual)

train$GarageCond <- as.character(train$GarageCond)
train$GarageCond[is.na(train$GarageCond)] <- "none"
train$GarageCond <- as.factor(train$GarageCond)

train$PoolQC <- as.character(train$PoolQC)
train$PoolQC[is.na(train$PoolQC)] <- "none"
train$PoolQC <- as.factor(train$PoolQC)

train$Fence <- as.character(train$Fence)
train$Fence[is.na(train$Fence)] <- "none"
train$Fence <- as.factor(train$Fence)

train$MiscFeature <- as.character(train$MiscFeature)
train$MiscFeature[is.na(train$MiscFeature)] <- "none"
train$MiscFeature <- as.factor(train$MiscFeature)

train$LotFrontage[is.na(train$LotFrontage)] <- 0

train <- train %>% filter(!is.na(Electrical) & !is.na(GarageYrBlt))

#Removing rare factors
train <- train %>% group_by(MSSubClass) %>% filter(n() >= 6) %>% 
  group_by(MSZoning) %>% filter(n() >= 6) %>%
  group_by(Alley) %>% filter(n() >= 6) %>%
  group_by(LotShape) %>% filter(n() >= 6) %>%
  group_by(LandContour) %>% filter(n() >= 6) %>%
  group_by(LotConfig) %>% filter(n() >= 6) %>%
  group_by(Neighborhood) %>% filter(n() >= 6) %>%
  group_by(Condition1) %>% filter(n() >= 6) %>%
  group_by(Condition2) %>% filter(n() >= 6) %>%
  group_by(BldgType) %>% filter(n() >= 6) %>%
  group_by(HouseStyle) %>% filter(n() >= 6) %>%
  group_by(OverallQual) %>% filter(n() >= 6) %>%
  group_by(OverallCond) %>% filter(n() >= 6) %>%
  group_by(RoofStyle) %>% filter(n() >= 6) %>%
  group_by(RoofMatl) %>% filter(n() >= 6) %>%
  group_by(Exterior1st) %>% filter(n() >= 6) %>%
  group_by(Exterior2nd) %>% filter(n() >= 6) %>%
  group_by(MasVnrType) %>% filter(n() >= 6) %>%
  group_by(ExterQual) %>% filter(n() >= 6) %>%
  group_by(ExterCond) %>% filter(n() >= 6) %>%
  group_by(Foundation) %>% filter(n() >= 6) %>%
  group_by(BsmtQual) %>% filter(n() >= 6) %>%
  group_by(BsmtCond) %>% filter(n() >= 6) %>%
  group_by(BsmtExposure) %>% filter(n() >= 6) %>%
  group_by(BsmtFinType1) %>% filter(n() >= 6) %>%
  group_by(BsmtFinType2) %>% filter(n() >= 6) %>%
  group_by(Heating) %>% filter(n() >= 6) %>%
  group_by(HeatingQC) %>% filter(n() >= 6) %>%
  group_by(CentralAir) %>% filter(n() >= 6) %>%
  group_by(KitchenQual) %>% filter(n() >= 6) %>%
  group_by(Functional) %>% filter(n() >= 6) %>%
  group_by(FireplaceQu) %>% filter(n() >= 6) %>%
  group_by(GarageType) %>% filter(n() >= 6) %>%
  group_by(GarageFinish) %>% filter(n() >= 6) %>%
  group_by(GarageQual) %>% filter(n() >= 6) %>%
  group_by(GarageCond) %>% filter(n() >= 6) %>%
  group_by(PavedDrive) %>% filter(n() >= 6) %>%
  group_by(PoolQC) %>% filter(n() >= 6) %>%
  group_by(Fence) %>% filter(n() >= 6) %>%
  group_by(SaleType) %>% filter(n() >= 6) %>%
  group_by(SaleCondition) %>% filter(n() >= 6)
  

train$MSSubClass <- factor(train$MSSubClass) 
train$MSZoning <- factor(train$MSZoning)
train$Alley <- factor(train$Alley)
train$LotShape <- factor(train$LotShape)
train$LandContour <- factor(train$LandContour)
train$LotConfig <- factor(train$LotConfig)
train$Neighborhood <- factor(train$Neighborhood)
train$Condition1 <- factor(train$Condition1)
train$Condition2 <- factor(train$Condition2)
train$BldgType <- factor(train$BldgType)
train$HouseStyle <- factor(train$HouseStyle)
train$OverallQual <- factor(train$OverallQual)
train$OverallCond <- factor(train$OverallCond)
train$RoofStyle <- factor(train$RoofStyle)
train$RoofMatl <- factor(train$RoofMatl)
train$Exterior1st <- factor(train$Exterior1st)
train$Exterior2nd <- factor(train$Exterior2nd)
train$MasVnrType <- factor(train$MasVnrType)
train$ExterQual <- factor(train$ExterQual)
train$ExterCond <- factor(train$ExterCond)
train$Foundation <- factor(train$Foundation)
train$BsmtQual <- factor(train$BsmtQual)
train$BsmtCond <- factor(train$BsmtCond)
train$BsmtExposure <- factor(train$BsmtExposure)
train$BsmtFinType1 <- factor(train$BsmtFinType1)
train$BsmtFinType2 <- factor(train$BsmtFinType2)
train$Heating <- factor(train$Heating)
train$HeatingQC <- factor(train$HeatingQC)
train$CentralAir <- factor(train$CentralAir)
train$KitchenQual <- factor(train$KitchenQual)
train$Functional <- factor(train$Functional)
train$FireplaceQu <- factor(train$FireplaceQu)
train$GarageType <- factor(train$GarageType)
train$GarageFinish <- factor(train$GarageFinish)
train$GarageQual <- factor(train$GarageQual)
train$GarageCond <- factor(train$GarageCond)
train$PavedDrive <- factor(train$PavedDrive)
train$PoolQC <- factor(train$PoolQC)
train$Fence <- factor(train$Fence)
train$SaleType <- factor(train$SaleType)
train$SaleCondition <- factor(train$SaleCondition)

```

Big question: Do different neighborhoods affect the housing prices?

```{r}

#Split data into 2006-2007 and 2008-2010
# train.precrash <- train %>% filter(YrSold <= 2007)
# train.postcrash <- train %>% filter(YrSold > 2007)
# 
# train.precrash <- train.precrash %>% dplyr::select(-YrSold)
# train.postcrash <- train.postcrash %>% dplyr::select(-YrSold)

#NEIGHBORING
table(train$Neighborhood)

library(ggplot2)

ggplot(train, aes(x = train$SalePrice, y = train$Neighborhood)) + geom_point(aes(colour = factor(train$Neighborhood))) + xlab("Sales Price") + 
  ylab("different neighborhood") + ggtitle("Impact of different neighborhood on sales price") + scale_colour_discrete("Neighborhood in Ames,IA")

#Checked the geometric characteristics for each neighborhoods from the map. Two groups were seperated from north and south

trian.north <- train %>% dplyr::filter(Neighborhood=="Blmngtn" |Neighborhood=="BrkSide"|Neighborhood=="Gilbert"|Neighborhood=="NAmes"||Neighborhood=="NoRidge"|Neighborhood=="NPkVill"|Neighborhood=="NridgHt"|Neighborhood=="BrDale"| Neighborhood=="NWAmes"| Neighborhood=="OldTown" |Neighborhood=="Somerst",Neighborhood=="StoneBr")

trian.south <- train %>% dplyr::filter(Neighborhood=="Bluestem"|Neighborhood=="ClearCr"|Neighborhood=="CollgCr"| Neighborhood=="Crawfor"| Neighborhood=="Edwards"|Neighborhood=="MeadowV"|  Neighborhood=="Mitchel"|Neighborhood=="SWISU"| Neighborhood=="Sawyer" |Neighborhood=="SawyerW" |  Neighborhood=="Timber"| Neighborhood=="Veenker")


#Scatter plot

library(ggplot2)

train$Neighborhoodcategory<- 
  ifelse(train$Neighborhood=="Blmngtn" |train$Neighborhood=="BrkSide"|train$Neighborhood=="Gilbert"|train$Neighborhood=="NAmes"|train$Neighborhood=="NoRidge"|train$Neighborhood=="NPkVill"|train$Neighborhood=="NridgHt"|train$Neighborhood=="BrDale"| train$Neighborhood=="NWAmes"| train$Neighborhood=="OldTown" |train$Neighborhood=="Somerst"|train$Neighborhood=="StoneBr", "1_North",
         ifelse(train$Neighborhood=="Bluestem"|train$Neighborhood=="ClearCr"|train$Neighborhood=="CollgCr"| train$Neighborhood=="Crawfor"|train$Neighborhood=="Edwards"|train$Neighborhood=="MeadowV"|  train$Neighborhood=="Mitchel"|train$Neighborhood=="SWISU"| train$Neighborhood=="Sawyer" |train$Neighborhood=="SawyerW" |  train$Neighborhood=="Timber"| train$Neighborhood=="Veenker", "2_South", "Other")
         )

ggplot(train, aes(x = train$SalePrice, y = train$Neighborhood)) + geom_point(aes(colour = factor(train$Neighborhoodcategory))) + xlab("Sales Price") + 
  ylab("different neighborhood") + ggtitle("Impact of different neighborhood on sales price") + scale_colour_discrete("Neighborhood in Ames,IA")

#Assumption: North is more expensive?

```

Select predictors that are available to the buyers on 'Zillow' website
```{r}

# subset of variables we want in our first model
train_info <- subset(train, select=c(
"LotArea",
"Neighborhood",
"YearBuilt",
"YearRemodAdd",
"TotalBsmtSF",
"GrLivArea",
"FullBath",
"HalfBath",
"BedroomAbvGr",
"Fireplaces",
"GarageCars",
"GarageArea",
"MoSold",
"YrSold",
"SalePrice","HouseStyle","SaleCondition"))

str(train_info)
```



```{r}

library(rpart)
library(rpart.plot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RCurl)
library(partykit)

require(rpart)
require(rpart.plot)

tree <-rpart( SalePrice ~., train_info, minbucket =1) 
prp(tree,nn.cex=1) 
```

The importance score for each variable can be obtained from the tree. GarageCars and neighborhood has the 1st and 2nd largest importance score among all variables. 
```{r}
print(tree$variable.importance)
```

Let's prune the tree 

```{r}
tree_0.05 <- prune(tree,cp=0.05)
# rpart.plot(tree_0.05,fallen.leaves=FALSE)
prp(tree_0.05,nn.cex=1)
```


Now we can check if there are some interaction terms in a linear regression model. The tree model used GarageCars as the first variable to split the model, which may indicate that this variable has the most non-linear relationship with the saleprice, our outcome. 

```{r}
library(ggplot2)

train_info$Neighborhoodcategory<- 
  ifelse(train$Neighborhood=="Blmngtn" |train$Neighborhood=="BrkSide"|train$Neighborhood=="Gilbert"|train$Neighborhood=="NAmes"|train$Neighborhood=="NoRidge"|train$Neighborhood=="NPkVill"|train$Neighborhood=="NridgHt"|train$Neighborhood=="BrDale"| train$Neighborhood=="NWAmes"| train$Neighborhood=="OldTown" |train$Neighborhood=="Somerst"|train$Neighborhood=="StoneBr", "1_North",
         ifelse(train$Neighborhood=="Bluestem"|train$Neighborhood=="ClearCr"|train$Neighborhood=="CollgCr"| train$Neighborhood=="Crawfor"|train$Neighborhood=="Edwards"|train$Neighborhood=="MeadowV"|  train$Neighborhood=="Mitchel"|train$Neighborhood=="SWISU"| train$Neighborhood=="Sawyer" |train$Neighborhood=="SawyerW" |  train$Neighborhood=="Timber"| train$Neighborhood=="Veenker", "2_South", "Other")
         )

#visualization of finding GarageCars^2 terms as one interaction term
p<-ggplot(train_info, aes(x = train_info$GarageCars, y = train_info$SalePrice),group=1) 
p<- p+stat_summary(fun.y=mean, colour="red", geom="line", size = 3)
p<- p + geom_point(aes(colour=train_info$Neighborhoodcategory)) 
p<- p+xlab("Numberof Garage") 
p<- p +ylab("Sales Price") 
p<- p+ggtitle("Impact of size of garage in car capacity on sales price") 
p<- p+scale_colour_discrete("Neighborhood in Ames,IA")
print(p)



```

Need to convert factor to integer -> extra columns needs to be made for
Neighborhood,HouseStyle,SaleCondition
```{r}

#Neighborhood
train_info$Neighborhood_North<- 
  ifelse(train$Neighborhood=="Blmngtn" |train$Neighborhood=="BrkSide"|train$Neighborhood=="Gilbert"|train$Neighborhood=="NAmes"|train$Neighborhood=="NoRidge"|train$Neighborhood=="NPkVill"|train$Neighborhood=="NridgHt"|train$Neighborhood=="BrDale"| train$Neighborhood=="NWAmes"| train$Neighborhood=="OldTown" |train$Neighborhood=="Somerst"|train$Neighborhood=="StoneBr", 1, 0)
         

train_info$Neighborhood_South<- ifelse(train$Neighborhood=="Bluestem"|train$Neighborhood=="ClearCr"|train$Neighborhood=="CollgCr"| train$Neighborhood=="Crawfor"|train$Neighborhood=="Edwards"|train$Neighborhood=="MeadowV"|  train$Neighborhood=="Mitchel"|train$Neighborhood=="SWISU"| train$Neighborhood=="Sawyer" |train$Neighborhood=="SawyerW" |  train$Neighborhood=="Timber"| train$Neighborhood=="Veenker", 1, 0)


#HouseStyle
train_info$HouseStyle_1<- 
  ifelse(train$HouseStyle=="1.5Fin" |train$HouseStyle=="1.5Unf"|train$HouseStyle=="1Story", 1, 0)

train_info$HouseStyle_2<- 
  ifelse(train$HouseStyle=="2.5Fin" |train$HouseStyle=="2.5Unf"|train$HouseStyle=="2Story", 1, 0)

train_info$HouseStyle_3<- 
  ifelse(train$HouseStyle=="SFoyer" |train$HouseStyle=="SLvl", 1, 0)

#SaleCondition

train_info$SaleCondition_normal<- 
  ifelse(train$SaleCondition=="Normal", 1, 0)


train_info$SaleCondition_notnormal<- 
  ifelse(train$SaleCondition=="Abnorml"|train$SaleCondition=="Alloca"|train$SaleCondition=="Family"|train$SaleCondition=="Partial", 1, 0)

str(train_info)

```


```{r}
# fit a full-scale model

lm.full <- lm(SalePrice ~ ., data = train)
summary(lm.full)

```


To build the accurate linear regression model, we checked correlinearity between variables
```{r}



```




```{r}
# Evaluate the variable importance by all subsets regression
# install.packages("leaps")
library(leaps)
leaps<-regsubsets(MMSCORE ~ ., data = AD,nbest=4)
# view results 
summary(leaps)
# plot a table of models showing variables in each model.
# models are ordered by the selection statistic.
plot(leaps,scale="r2")
```


```{r}
#clustering

require( 'ggplot2' )
p <- qplot( data=linear.data, x.1, x.2, colour=factor(SalePrice) )
p <- p + labs(title = "Scatterplot of data points of two classes")
print(p)
```
